---
title: "Final Report"
subtitle: "Deconvolution "
author: "Marina Vilardell"
date: "2023-08-16"
output: 

  BiocStyle::html_document:
    number_sections: true
    toc_float: yes
---

Before applying deconvolution, we have to make sure that the input data is: 

1. Normalized

2. not log-transformed.

The data was normalized in the previous steps, and the transformation of not log data will be performed next, when running the deconvolution. 
## Cibersort

In order to use Cibersort, we require to have in local the cibersort source code, that contains two files. These files are downloaded from the cibersort website. 

```{r}
#Set the path to the directory were the deconvolution cibersort files are. 
deconvDir <- file.path("C:/Users/MARINA/Documents/MASTER/TFM/Deconvolution_files")
cibersort_binary = file.path(deconvDir, "CIBERSORT.R")
cibersort_mat = file.path(deconvDir, "CIBERSORT_LM22.txt")

CIBERSORT_LM22 <- read.delim("~/MASTER/TFM/Deconvolution_files/CIBERSORT_LM22.txt", row.names=1)


# Separate radiated from non-radiated
NOR<-pdata[pdata$Treatment=='NOR',]
RAD<-pdata[pdata$Treatment=='RAD',]
# 
radiated_samples<-rownames(RAD)
non_radiated_samples<-rownames(NOR)


```

## Cibersort abs. mode

We will use cibersort abs. mode, which is a method that provides an score that allows for the comparison of both between cell types, and samples comparisons. 

```{r,warning=FALSE}
library(immunedeconv)
res_cib_abs=deconvolute_cibersort(2^exprs(brca.noduplicated),arrays = TRUE, absolute = TRUE)
#load(file='res_cis_abs.RData')

res_cib_abs <- data.frame(cell_type=rownames(res_cib_abs),res_cib_abs)
res_cib_abs <- res_cib_abs[,-1]
res_cib_abs <- res_cib_abs %>% mutate(across(where(is.numeric), round, 3))
colnames(res_cib_abs)<-gsub('_2','',colnames(res_cib_abs))
datatable(res_cib_abs) 


# ---  Load the data ---
library(readxl)
Patient_info <- as.data.frame(read_excel("../Patient info.xlsx"))

# --- Prepare it ---
# Change the row names for their array code identification
rownames(Patient_info)<-Patient_info$`ARRAY CODE`
rownames(Patient_info)<-gsub('LL','L',rownames(Patient_info))
rownames(Patient_info)<-gsub('SG-127.CEL','X127.CEL',rownames(Patient_info))
rownames(Patient_info)<-gsub('SG-13.CEL','SG-103.CEL',rownames(Patient_info))
rownames(Patient_info)<-gsub('SG-14.CEL','SG-104.CEL',rownames(Patient_info))
rownames(Patient_info)<-gsub('-','.',rownames(Patient_info))

ind<-match(colnames(res_cib_abs),row.names(Patient_info))
Patient_info<-Patient_info[ind,]
identical(colnames(res_cib_abs),row.names(Patient_info))

```

```{r}
# Select non-radiated samples
noradiated<-rownames(pdata[pdata$Treatment=='NOR',])
noradiated<-gsub('-','.',noradiated)
noradiated<-gsub('127.CEL','X127.CEL',noradiated)
res_cib_abs_norad<-res_cib_abs[,noradiated]

# Select radiated samples
radiated<-rownames(pdata[pdata$Treatment=='RAD',])
radiated<-gsub('-','.',radiated)
radiated<-gsub('SG.28_2.CEL','SG.28.CEL',radiated)
res_cib_abs_rad<-res_cib_abs[,radiated]

```


Once we have obtained the table of the cell fractions, we can start by assessing its variability in the different conditions we want to study. 

In order to know which statistical test is the most appropriate one, we will start by performing a shapiro.test, to assess the normality of the data. 

For each cell type, 2 hypothesis are made: 

- H0: Data follow a normal distribution.

- H1: Data do not follow a normal distribution.

If p-value is greater than 0.05, data follows a normal distribution, whereas if the p-value is less than 0.05 data do not follow a normal distribution, and this will condition the type of test to be used. 

```{r}
# remove cell types that do not have at least 5 unique values
distinct_values_count <- apply(res_cib_abs_norad, 1, function(x) length(unique(x)))

# Subset the data frame to include rows with at least 5 distinct values
df_filtered <- res_cib_abs_norad[distinct_values_count >= 5, ]


test<-apply(df_filtered,1,function(x) shapiro.test(x)$p.value)
test
notdistributed<-df_filtered

#Normal distribution
hist(as.numeric(df_filtered[16,]))

# Not normal distribution
hist(as.numeric(df_filtered[8,]))

notdistributed<-notdistributed[-18,]
notdistributed$cell_type<-rownames(notdistributed)

notdistributed %>%
  gather(sample, fraction, -cell_type) %>%
  ggplot(aes(x=sample, y=fraction, fill=cell_type)) +
    geom_bar(stat='identity') +
    scale_x_discrete(limits = rev(levels(notdistributed))) +
    theme(text = element_text(size = 10)) + theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))+ggtitle('Cell type fractions') + scale_fill_discrete(name = "Cell type")
```

As most of the data do not follow a normal distribution, we will consider all the cell types to be not normally distributed. Therefore, we will apply a Wilcoxon for testing paired data for testing the equality of 2 means, a Kruskall wallis test for the equality of more than 2 means. 

In this code, the mean for each cell type is calculated by each group. We will assess the variability of cell types depending the phenotype of the individual. 
```{r}
# vector of conditions
noradiated<-rownames(pdata[pdata$Treatment=='NOR',])
pdata_norad<-pdata[noradiated,]
condition_rad_norad=pdata_norad$Phenotype

notdistributed<-notdistributed[,-54]

mean_values<-NULL
for (i in 1: nrow(notdistributed)){
    mean<-tapply((unlist(notdistributed[i,])),condition_rad_norad, mean)
    rownam<-rownames(notdistributed[i,])
    mean_values<-c(mean_values,c(rownam,mean))
    }


# Function to format the mean_values into groups of 7
format_mean_values <- function(values) {
  result <- list()
  num_values <- length(values)
  i <- 1
  while (i <= num_values) {
    result <- c(result, list(values[i:(i + 6)]))
    i <- i + 7
  }
  return(result)
}

mean_values <- format_mean_values(mean_values)

# Convert mean_values to a data frame
mean_values_df <- do.call(rbind, mean_values)
mean_values_df <- as.data.frame(mean_values_df)

# Rename the columns
colnames(mean_values_df) <- c("Cell type", "BRCA1.AFFECTED", "BRCA1.HEALEHY",'BRCA2.AFFECTED','BRCA2.HEALTHY','NOMUT.AFFECTED','NOMUT.HEALTHY')

mean_values_df$BRCA1.AFFECTED<-round(as.numeric(mean_values_df$BRCA1.AFFECTED),4)
mean_values_df$BRCA1.HEALEHY<-round(as.numeric(mean_values_df$BRCA1.HEALEHY),4)
mean_values_df$BRCA2.AFFECTED<-round(as.numeric(mean_values_df$BRCA2.AFFECTED),4)
mean_values_df$BRCA2.HEALTHY<-round(as.numeric(mean_values_df$BRCA2.HEALTHY),4)
mean_values_df$NOMUT.AFFECTED<-round(as.numeric(mean_values_df$NOMUT.AFFECTED),4)
mean_values_df$NOMUT.HEALTHY<-round(as.numeric(mean_values_df$NOMUT.HEALTHY),4)
mean_values_df
```

Apply statistical tests:

```{r}
pvals<-NULL
for (i in 1:nrow(notdistributed)){
  pval<-kruskal.test(unlist(notdistributed[i,])~condition_rad_norad)$p.val
  pvals<-c(pvals,pval)
}

notdistributed$pvals<-pvals
notdistributed<-notdistributed[,-54]

df<-notdistributed[,106:107] #106:107
df[-1]

notdistributed$p.adjusted<-p.adjust(notdistributed$pvals, method = "bonferroni", n = length(notdistributed$pvals))

```


#### Total CD4 cell type

The main interest of this study is the T CD4 cell type. As cibersort provides detailed information regarding them, and separates the lymphocytes depending on their subgroups (naive or memory activated),we will aggregate them by summing up their expression value for each sample. 

```{r}
# Select T cells CD4 naive
T_CD4_naive_ciber<-notdistributed['T cells CD4 naive',]
T_CD4_naive_ciber<-as.data.frame(t(T_CD4_naive_ciber))


# Select T cells CD4 memory activated
T_CD4_memory_act_ciber<-notdistributed['T cells CD4 memory activated',]
T_CD4_memory_act_ciber<-as.data.frame(t(T_CD4_memory_act_ciber))

# Select total T CD4 cells 
TCD4_cibersort<-notdistributed[c('T cells CD4 naive','T cells CD4 memory activated'),]

TCD4_cibersort<-TCD4_cibersort[,-54]

Total_T_CD4_ciber<-colSums(TCD4_cibersort)
Total_T_CD4_ciber<-as.data.frame(Total_T_CD4_ciber)
Total_T_CD4_ciber

```

### Boxplot

```{r}
library(ggplot2)

noradiated<-rownames(pdata[pdata$Treatment=='NOR',])
noradiated2<-as.vector(gsub('-','.',noradiated))
noradiated2<-as.vector(gsub('127.CEL','X127.CEL',noradiated2))

res_norad<-notdistributed[,noradiated2]
sub_resnorad<-res_norad[c('T cells CD4 naive','T cells CD4 memory activated'),]
rownames(sub_resnorad)[2]<-'TCD4 memory activated'

#----------------BRCA1 affected-------------------------------------
BRCA1.AF<-rownames(pdata_norad[pdata_norad$Phenotype=='BRCA1.AF',])
BRCA1.AF<-as.vector(gsub('-','.',BRCA1.AF))
pos<-which(colnames(sub_resnorad) %in% BRCA1.AF)

res_noradBRCA1.AF<-sub_resnorad[,pos]

# Put all data in a single vector
vectors<-NULL
for (i in 1:ncol(res_noradBRCA1.AF)){
  vector<-res_noradBRCA1.AF[,i]
  vectors<-c(vectors,vector)
}

data_plot <- data.frame(
  phenotype = rep('BRCA1.AFFECTED', each = 2),
  cell_type = rep(rownames(res_noradBRCA1.AF), times = 10),
  value=vectors)

#----------------BRCA2 affected-------------------------------------
BRCA2.AF<-rownames(pdata_norad[pdata_norad$Phenotype=='BRCA2.AF',])
BRCA2.AF<-as.vector(gsub('-','.',BRCA2.AF))
pos<-which(colnames(sub_resnorad) %in% BRCA2.AF)

res_noradBRCA2.AF<-sub_resnorad[,pos]

# Put all data in a single vector
vectors<-NULL
for (i in 1:ncol(res_noradBRCA2.AF)){
  vector<-res_noradBRCA2.AF[,i]
  vectors<-c(vectors,vector)
}

data_plot_BRCA2.AF <- data.frame(
  phenotype = rep('BRCA2.AFFECTED', each = 2),
  cell_type = rep(rownames(res_noradBRCA2.AF), times = 11),
  value=vectors)


#----------------BRCA1 healthy-------------------------------------
BRCA1.HE<-rownames(pdata_norad[pdata_norad$Phenotype=='BRCA1.SA',])
BRCA1.HE<-as.vector(gsub('-','.',BRCA1.HE))
pos<-which(colnames(sub_resnorad) %in% BRCA1.HE)

res_noradBRCA1.HE<-sub_resnorad[,pos]

# Put all data in a single vector
vectors<-NULL
for (i in 1:ncol(res_noradBRCA1.HE)){
  vector<-res_noradBRCA1.HE[,i]
  vectors<-c(vectors,vector)
}

data_plot_BRCA1.HE <- data.frame(
  phenotype = rep('BRCA1.HEALTHY', each = 2),
  cell_type = rep(rownames(res_noradBRCA1.HE), times = 7),
  value=vectors)

#----------------BRCA2 healthy-------------------------------------
BRCA2.HE<-rownames(pdata_norad[pdata_norad$Phenotype=='BRCA2.SA',])
BRCA2.HE<-as.vector(gsub('-','.',BRCA2.HE))
pos<-which(colnames(sub_resnorad) %in% BRCA2.HE)

res_noradBRCA2.HE<-sub_resnorad[,pos]

# Put all data in a single vector
vectors<-NULL
for (i in 1:ncol(res_noradBRCA2.HE)){
  vector<-res_noradBRCA2.HE[,i]
  vectors<-c(vectors,vector)
}

data_plot_BRCA2.HE <- data.frame(
  phenotype = rep('BRCA2.HEALTHY', each = 2),
  cell_type = rep(rownames(res_noradBRCA2.HE), times = 9),
  value=vectors)


#----------------NOMUT affected-------------------------------------
NOMUT.AF<-rownames(pdata_norad[pdata_norad$Phenotype=='NOMUT.AF',])
NOMUT.AF<-as.vector(gsub('-','.',NOMUT.AF))
pos<-which(colnames(sub_resnorad) %in% NOMUT.AF)

res_noradNOMUT.AF<-sub_resnorad[,pos]

# Put all data in a single vector
vectors<-NULL
for (i in 1:ncol(res_noradNOMUT.AF)){
  vector<-res_noradNOMUT.AF[,i]
  vectors<-c(vectors,vector)
}

data_plot_NOMUT.AF <- data.frame(
  phenotype = rep('NOMUT.AFFECTED', each = 2),
  cell_type = rep(rownames(res_noradNOMUT.AF), times = 5),
  value=vectors)

#----------------NOMUT healthy-------------------------------------
NOMUT.HE<-rownames(pdata_norad[pdata_norad$Phenotype=='NOMUT.SA',])
NOMUT.HE<-as.vector(gsub('-','.',NOMUT.HE))
pos<-which(colnames(sub_resnorad) %in% NOMUT.HE)

res_noradNOMUT.HE<-sub_resnorad[,pos]

# Put all data in a single vector
vectors<-NULL
for (i in 1:ncol(res_noradNOMUT.HE)){
  vector<-res_noradNOMUT.HE[,i]
  vectors<-c(vectors,vector)
}

data_plot_NOMUT.HE <- data.frame(
  phenotype = rep('NOMUT.HEALTHY', each = 2),
  cell_type = rep(rownames(res_noradNOMUT.HE), times = 10),
  value=vectors)

data_final<-rbind(data_plot,data_plot_BRCA1.HE,data_plot_BRCA2.AF,data_plot_BRCA2.HE,data_plot_NOMUT.AF,data_plot_NOMUT.HE)

library(rstatix)
stat.test<-data_final %>% group_by(cell_type) %>% kruskal_test(value ~ phenotype)


stat.test2<-data_final %>% group_by(cell_type) %>% wilcox_test(value ~ phenotype)


plot<-ggplot(data_final, aes(x=phenotype, y=value, fill=cell_type)) + 
  geom_boxplot(outlier.shape = NA)+facet_wrap(~cell_type) +theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) + ggtitle('Distribution of T cells CD4 per phenotype')

data_text <- data.frame(label = c("p=0.12", "p=0.08"),  # Create data for text
                        cell_type = c('T cells CD4 naive','TCD4 memory activated'),
                        x = c(2,2),
                        y = c(1,1))

plot + geom_text(data=data_text, mapping=aes(x=x,y=y,label=label))  + scale_fill_discrete(name = "Cell type")
```



### Extra analysis

As we had all the data available, including irradiated samples, researches asked to assess the differences of cell type expression levels between non-irradiated and irradiated samples. 

We will start by a general overview, so we will assess the mean differences between 2 general groups: radiated vs non-radiated samples.  

In this code, the mean for each cell type is calculated by each group.

```{r}
# remove cell types that do not have at least 5 unique values
distinct_values_count <- apply(res_cib_abs, 1, function(x) length(unique(x)))

# Subset the data frame to include rows with at least 5 distinct values
df_filtered <- res_cib_abs[distinct_values_count >= 5, ]


test<-apply(df_filtered,1,function(x) shapiro.test(x)$p.value)
test
notdistributed_all<-df_filtered

# Normal distribution 
hist(as.numeric(df_filtered[16,]))

# Not normal distribution
hist(as.numeric(df_filtered[8,]))

# Separate radiated from non-radiated
table(pdata$Treatment)

# vector of conditions
condition_rad_norad=pdata$Treatment

mean_values<-NULL
for (i in 1: nrow(notdistributed_all)){
    mean<-tapply((unlist(notdistributed_all[i,])),condition_rad_norad, mean)
    rownam<-rownames(notdistributed_all[i,])
    mean_values<-c(mean_values,c(rownam,mean))
    }

# Function to format the mean_values into groups of 3
format_mean_values <- function(values) {
  result <- list()
  num_values <- length(values)
  i <- 1
  while (i <= num_values) {
    result <- c(result, list(values[i:(i + 2)]))
    i <- i + 3
  }
  return(result)
}

mean_values <- format_mean_values(mean_values)

# Convert mean_values to a data frame
mean_values_df <- do.call(rbind, mean_values)
mean_values_df <- as.data.frame(mean_values_df)

# Rename the columns
colnames(mean_values_df) <- c("Condition", "NOR", "RAD")

mean_values_df$NOR<-round(as.numeric(mean_values_df$NOR),4)
mean_values_df$RAD<-round(as.numeric(mean_values_df$RAD),4)

mean_values_df
```

As you can see there is no big differences between the mean cell fraction of no radiated and radiated samples, and not in all cell types the mean is higher in radiated samples rather the no radiated (ex: T cells CD4 memory activated). However, we will perform a wilcoxon test to assess if there are statistical differences. 

Apply statistical tests:
```{r}
# --- radiated vs no-radiated --------------

pvals<-NULL
for (i in 1:nrow(notdistributed_all)){
  pval<-wilcox.test(unlist(notdistributed_all[i,])~condition_rad_norad)$p.val
  pvals<-c(pvals,pval)
}

notdistributed_all$pvals<-pvals

df<-notdistributed_all[,106:107]
df<-df[-1]

df$ad.pval<-p.adjust(df$pvals, method = "bonferroni", n = length(df$pvals))
df
```

There are 2 p-values that are less than 0.05, which means that there are statistically evidences to reject the null hypothesis (H0) and thus, the alternative hypothesis (H1) is considered to be proved. So, our results provide support for the hypothesis that there are significant differences between the mean expression of the radiated and non radiated levels of T cells regulatory (Tregs), and Macrophages MO. 

## EPIC

We will also perform deconvolution using the EPIC method. 

```{r,warning=FALSE}
res_epic2 = deconvolute(2^exprs(brca.noduplicated),array=TRUE , "epic") 
res_epic <- res_epic2 %>% mutate(across(where(is.numeric), round, 3))
datatable(res_epic) 

res_epic<-as.data.frame(res_epic)
rownames(res_epic)<-res_epic$cell_type
res_epic<-res_epic[,-1]


noradiated<-rownames(pdata[pdata$Treatment=='NOR',])
res_epic_norad<-res_epic[,noradiated]

test<-apply(res_epic_norad,1,function(x) shapiro.test(x)$p.value)
test
pos<-which(test<0.05)
notdistributed<-res_epic[pos,]


```

### T cell CD4 EPIC

Let's prepare the T cell CD4 variable to perform a correlation analysis. 
```{r}
# Get T CD4+ cells expression levels. 
T_CD4_epic<-res_epic_norad['T cell CD4+',]
T_CD4_epic<-t(T_CD4_epic)
T_CD4_epic<-as.data.frame(T_CD4_epic)
T_CD4_epic
```



## MIXTURE

Finally, we will also perform deconvolution by using the MIXTURE method, based in cibersort. 

```{r,warning=FALSE,message=FALSE}
library(MIXTURE)
res_mixt<-MIXTURE(2^exprs(brca.noduplicated),signatureMatrix = CIBERSORT_LM22)
res_m<-as.data.frame(t(res_mixt$Subjects$MIXabs))

res_m_norad<-res_m[,noradiated]

# remove cell types that do not have at least 5 unique values
distinct_values_count <- apply(res_m_norad, 1, function(x) length(unique(x)))

# Subset the data frame to include rows with at least 5 distinct values
df_filtered <- res_m_norad[distinct_values_count >= 5, ]

```


#### Total T CD4 cells

```{r}
# Select T cells CD4 naive
T_CD4_naive_mixt<-df_filtered['T.cells.CD4.naive',]
T_CD4_naive_mixt<-as.data.frame(t(T_CD4_naive_mixt))


# Select T cells CD4 memory activated
T_CD4_memory_act_mixt<-df_filtered['T.cells.CD4.memory.activated',]
T_CD4_memory_act_mixt<-as.data.frame(t(T_CD4_memory_act_mixt))


# Select T CD4 cells 
TCD4_mixture<-df_filtered[c('T.cells.CD4.naive','T.cells.CD4.memory.activated'),]

#apply(TCD4_mixture,2,sum)
Total_T_CD4_mixture<-colSums(TCD4_mixture)
Total_T_CD4_mixture<-as.data.frame(Total_T_CD4_mixture)
Total_T_CD4_mixture
```


In conclusion from the deconvolution analysis, the proportion of T CD4 cell type in all the 3 methods is very high. 

## CORRELATIONS

In this section, correlation analysis between the 3 deconvolution methods used will be performed. The goal is to try to identify which method will be used to obtain the TCD4 cell type covariate to include it in the variable selection with LASSO step. 

### Correlation EPIC-MIXTURE

```{r,warning=FALSE,message=FALSE}
library(ggplot2)
library(ggpubr)
T_cellcd4_mixt_epic<-cbind(Total_T_CD4_mixture,T_CD4_epic)
colnames(T_cellcd4_mixt_epic)<-c('TCD4_MIXTURE','TCD4_EPIC')

cor.test(T_cellcd4_mixt_epic$TCD4_MIXTURE,T_cellcd4_mixt_epic$TCD4_EPIC,method = 'spearman')

ggplot(data=T_cellcd4_mixt_epic,aes(x = TCD4_MIXTURE, y = TCD4_EPIC)) + geom_point(color='black')+stat_smooth(method = 'lm') + 
  ggtitle('Correlation TCD4 EPIC and TCD4 MIXTURE') +xlab('TCD4 MIXTURE') +ylab('TCD4 EPIC') + geom_text(x=0.25,y=0.92,label='Rho = 0.51')
```

### Correlation CIBERSORT-MIXTURE

```{r,warning=FALSE}
T_cellcd4_ciber_mixt<-cbind(Total_T_CD4_ciber,Total_T_CD4_mixture)
colnames(T_cellcd4_ciber_mixt)<-c('TCD4_CIBERSORT_abs','TCD4_MIXTURE')

cor.test(T_cellcd4_ciber_mixt$TCD4_CIBERSORT_abs,T_cellcd4_ciber_mixt$TCD4_MIXTURE,method = 'spearman')

ggplot(data=T_cellcd4_ciber_mixt,aes(x = TCD4_CIBERSORT_abs, y = TCD4_MIXTURE)) + geom_point(color='black')+stat_smooth(method = 'lm') + 
  ggtitle('Correlation TCD4 CIBERSORT abs mode and TCD4 MIXTURE') +xlab('TCD4 CIBERSORT abs mode') +ylab('TCD4 MIXTURE') + geom_text(x=0.35,y=0.92,label='Rho = 0.86')
```


As both Cibersort and MIXTURE present detailed TCD4 cell types, we will assess the correlation between TCD4 naive and memory activated cell types. 

#### CD4 NAIVE

```{r,warning=FALSE}
T_cd4_naive_both<-cbind(T_CD4_naive_ciber,T_CD4_naive_mixt)
colnames(T_cd4_naive_both)<-c('naive_CIBERSORT_abs','naive_MIXT')

cor.test(T_cd4_naive_both$naive_CIBERSORT_abs,T_cd4_naive_both$naive_MIXT,method = 'spearman')

ggplot(data=T_cd4_naive_both,aes(x = naive_CIBERSORT_abs, y = naive_MIXT)) + geom_point(color='black')+stat_smooth(method = 'lm') + 
  ggtitle('Correlation TCD4 naive CIBERSORT abs mode and TCD4 naive MIXTURE') +xlab('TCD4 naive CIBERSORT abs mode') +ylab('TCD4 naive MIXTURE')
```

#### CD4 MEMORY ACTIVATED

```{r,warning=FALSE}
T_cd4_MEMORY_both<-cbind(T_CD4_memory_act_ciber,T_CD4_memory_act_mixt)
colnames(T_cd4_MEMORY_both)<-c('CIBERSORT_abs','MIXT')

cor.test(T_cd4_MEMORY_both$CIBERSORT_abs,T_cd4_MEMORY_both$MIXT,method = 'spearman')

ggplot(data=T_cd4_MEMORY_both,aes(x = CIBERSORT_abs, y = MIXT)) + geom_point(color='black')+stat_smooth(method = 'lm') + 
  ggtitle('Correlation TCD4 memory activated CIBERSORT and  MIXTURE') +xlab('TCD4 memory activated CIBERSORT abs mode') +ylab('TCD4 memory activated MIXTURE')
```

### Correlation CIBERSORT-EPIC

```{r,warning=FALSE}
T_cellcd4_ciber_epic<-cbind(Total_T_CD4_ciber,T_CD4_epic)
colnames(T_cellcd4_ciber_epic)<-c('TCD4_CIBERSORT_abs','TCD4_EPIC')

cor.test(T_cellcd4_ciber_epic$TCD4_CIBERSORT_abs,T_cellcd4_ciber_epic$TCD4_EPIC,method = 'spearman')

ggplot(data=T_cellcd4_ciber_epic,aes(x = TCD4_CIBERSORT_abs, y = TCD4_EPIC)) + geom_point(color='black')+stat_smooth(method = 'lm') + 
  ggtitle('Correlation TCD4 CIBERSORT abs mode and TCD4 EPIC') +xlab('TCD4 CIBERSORT') +ylab('TCD4 EPIC') + geom_text(x=0.35,y=0.92,label='Rho = 0.40')
```